---
title: "FinancialReturns"
author: "Eric He"
date: "July 17, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("dplyr")
library("lubridate")
library("tidyr")
library("purrr")
```

Financial returns data was downloaded from the Center for Research in Security Prices (CRSP) daily stock returns database from the Wharton Research Data Services (WRDS) account. Data from 01/01/2000 to 12/31/2016 was downloaded; every single ticker and data category was downloaded, just in case. Date was selected to be in MM/DD/YYYY form.

Market cap data was also downloaded from the CRSP database provided by WRDS, same beginning and end date, every single ticker and data category again. Date is in YYYYMMDD form.

Load in the data.

```{r}
master_index <- read.csv("masterIndex.csv")
full_returns <- read.csv("../Data/Financial Data/Trimmed_Returns_Raw.csv")
full_cap <- read.csv("market_cap.csv") %>%
  select(permno = PERMNO, num_shares = SHROUT, date = SHRSDT)
```

```{r}
head(masterIndex)
```

```{r}
head(full_returns)
```

```{r}
head(full_cap)
```

Select the relevant columns: holding period returns (RET), tickers (TICKER), company name (COMNAM), delisting return (DLRET), Shares Observation End Date (shrenddt)

End product: correct monthly returns

```{r}
full_returns <- select(full_returns, permco = PERMCO, date = date, return = RET, ticker = TICKER, delisting_return = DLRET) %>%
  filter(ticker %in% master_index$TICKER) %>% # only need tickers for which we have filings to compare with
  mutate(delisting_return = as.numeric(as.character(delisting_return))) %>% # 
  mutate(delisting_return = replace(delisting_return, is.na(delisting_return) == TRUE, 0)) %>% # replace NA with 0
  mutate(delisting_return = delisting_return + 1) %>% # so we can add 1 so we can multiply
  mutate(return = as.numeric(as.character(return))) %>% # change return from factor to numeric, characters which CRSP uses to represent missing data or point towards delisting return gets changed to NA
  mutate(return = replace(return, is.na(return) == TRUE, 0)) %>%
  mutate(return = return + 1) %>%
  filter(is.na(return) == FALSE) %>%
  mutate(return = return * delisting_return) %>%
  mutate(date = mdy(date)) %>%
  select(-delisting_return)
```

Given a ticker and a date from the master Index, devise a formula to calculate holding period returns for variable time intervals.

```{r}
the_ticker <- "AAPL"
the_year <- 2013
the_month <- 10
the_day <- 30

interval_length <- 3
interval_type <- "month"

start_date <- ymd("20131030")

end_date <- the_date %m+% months(3)

returns_calculator <- function(the_ticker, filing_date, interval_length = 5, interval_type = "day"){
  start_date_beginning <- filing_date %>% # start date is first non-weekend day before the filing date.
    ymd(.) %m-% days(3)
  start_date_candidates <- seq(start_date_beginning, ymd(filing_date) %m-% days(1), by = "days")
  start_date <- start_date_candidates[which(format(start_date_candidates, "%u") %in% c(1:5))[length(which(format(start_date_candidates, "%u") %in% c(1:5)))]] # pick out first weekday in the sequence, %u formats date object into numeric weekday
  end_date <- filing_date %>%
    ymd %>% # have to ram through ymd again because map() uses [[]] which messes with lubridate object type
    when(interval_type == "day" ~ . %m+% days(interval_length),
         interval_type == "month" ~ . %m+% months(interval_length), 
         interval_type == "year" ~ . %m+% years(interval_length),
         ~ stop(print(interval_type)))
  date_sequence <- seq(start_date, end_date, by = "days") # see above comment
  date_returns <- filter(full_returns, as.character(ticker) == as.character(the_ticker), date %in% date_sequence) # tickers must be converted to character or else throws a factor level error
  if (nrow(date_returns) == 0){
    empty_df <- data_frame(hpr = NA)
    print(paste("No financial data for ticker", the_ticker))
    return(empty_df) # if no financial data then we would like to make that clear
  } # when statement does not work to break the function! :(
  hpr <-  summarise(date_returns, hpr = prod(return))
  print(paste("Calculated hpr for ticker", the_ticker, "and date", filing_date))
  return(hpr)}
```

```{r}
returns_5_days <- map2_df(masterIndex$TICKER, masterIndex$DATE_FILED, returns_calculator)
returns_1_month <- map2_df(masterIndex$TICKER, masterIndex$DATE_FILED, returns_calculator, interval_length = 1, interval_type = "month")
returns_3_months <- map2_df(masterIndex$TICKER, masterIndex$DATE_FILED, returns_calculator, interval_length = 3, interval_type = "month")
returns_6_months <- map2_df(masterIndex$TICKER, masterIndex$DATE_FILED, returns_calculator, interval_length = 6, interval_type = "month")
returns_1_year <- map2_df(masterIndex$TICKER, masterIndex$DATE_FILED, returns_calculator, interval_length = 1, interval_type = "year")
```

The volatility calculator uses the original trimmed raw returns data to calculate, since it is more accurate about which days are trading days!

```{r}
volatility_calculator <- function(the_ticker, filing_date, interval_length = 1, interval_type = "year"){
  end_date <- filing_date %>% ymd
  start_date <- end_date %>%
    ymd %>%
    when(interval_type == "day" ~ . %m-% days(interval_length),
         interval_type == "month" ~ . %m-% months(interval_length), 
         interval_type == "year" ~ . %m-% years(interval_length),
         ~ stop(print(interval_type)))
  date_sequence <- seq(start_date, end_date, by = "days")
  date_sequence <- date_sequence[-which(format(date_sequence, "%u") %in% c(6,7))]
  date_returns <- filter(full_returns, as.character(ticker) == as.character(the_ticker), date %in% date_sequence)
  if (nrow(date_returns) == 0){
    empty_df <- data_frame(sd = NA)
    print(paste("No financial data for ticker", the_ticker))
    return(empty_df) 
  }
  volatility <-  summarise(date_returns, sd = sd(return))
  print(paste("Calculated volatility for ticker", the_ticker, "and date", filing_date))
  return(volatility)
}
```

```{r}
vol_1year <- map2_df(master_index$TICKER, master_index$DATE_FILED, volatility_calculator)
```

```{r}
master_index <- cbind(master_index, vol_1year)
master_index %<>% mutate(
  adj_ret5d = ret5d / (sqrt(5) * sd),
  adj_ret1m = ret1m / (sqrt(30) * sd),
  adj_ret3m = ret3m / (sqrt(90) * sd),
  adj_ret1y = ret1y / (sqrt(365) * sd)
)
```

```{r}
masterIndex <- cbind(masterIndex, ret5d = returns_5_days$hpr, ret1m = returns_1_month$hpr, ret3m = returns_3_months$hpr,  ret6m = returns_6_months$hpr, ret1y = returns_1_year$hpr)
write.csv(master_index, "../Data/Master Index/masterIndex.csv", row.names = FALSE)
```

