---
title: "sentimentAnalysisAlgo"
author: "Eric He"
date: "August 7, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library("quanteda")
library("dplyr")
library("purrr")
library("ggplot2")
library("reshape2")
library("gridExtra")
```

Read in the data relevant for all four years.

The desired sentiment analysis algorithm computes sentiment scores of every 

```{r}
negative <- readLines("negative.txt")
positive <- readLines("positive.txt")
sections <- c("1", "1A", "1B", "2", "3", "4", "5", "6", "7", "7A", "8", "9", "9A", "9B", "10", "11", "12", "13", "14", "15")
masterIndex <- read.csv("masterIndex.csv")
tickers <- unique(masterIndex$TICKER)
```









Build two functions, one which subsets the bigdfm according to section and the other recording the indices if the subsetted.

```{r}
indices_tfidf <- function(meta_df, cond){
  index_num <- filter(meta_df, section == cond) %>%
    select(filing)
  return(index_num)
}
subset_tfidf <- function(meta_df, dfmobj, cond){
  index_num <- filter(meta_df, section == cond) %>%
    select(index)
  weightsdfm <- dfmobj[index_num$index,] %>%
    tfidf(scheme_tf = "logave")
  return(weightsdfm)
}
```

Build a function which computes the sentiment scores. Recall that the sentiment score is the weights of the words in the sentiment dictionary, divided by the total weight of the words in the document.

```{r}
dfmstat_ratio <- function(dfmObj, dict){
  dfm_select(dfmObj, pattern = dict) %>%
    rowSums(.) / rowSums(dfmObj)
}
```

We would like to do the algorithm for the positive, negative, positive-negative sentiment scorings. This requires the positive and negative dictionaries.
Then we would like to do them for all four years. 

The indices_list sorts the metadata frame by section. 

```{r}
indices_list <- map(sections, indices_tfidf, meta_df = metadata)
weightsdfm_list <- map(sections, subset_tfidf, meta_df = metadata, dfmobj = bigdfm)
sentiment_list <- weightsdfm_list %>%
  map(dfmstat_ratio, dict = negative)
```

Now do this for every year's worth of data. The weighter function is used to save the indices and weightsdfm for every section in a given year. It is mapped across all four years of data.

```{r}
years <- c(2013:2016)
path_to_metadata <- "metadata"
metadata_type <- ".csv"
path_to_parsedDFM <- "parsedBigDfm"
parsedDFM_type <- ".RData"

weighter <- function(year){
metadata <- paste(path_to_metadata, year, metadata_type, sep = "") %>%
  read.csv()
load(paste(path_to_parsedDFM, year, parsedDFM_type, sep = ""))

indices_list <- map(sections, indices_tfidf, meta_df = metadata)
weightsdfm_list <- map(sections, subset_tfidf, meta_df = metadata, dfmobj = bigdfm)

save(indices_list, file = paste("indices_list_", year, ".RData", sep = ""))
save(weightsdfm_list, file = paste("weightsdfm_list_", year, ".RData", sep = ""))
}

map(years, weighter) # The weighter function returns nothing which is fine.
```

Get every sentiment list for every year.

The returns quantiler splits the data of each section into five quantiles and calculates the mean returns for a portfolio which invests equally in all companies of the quantile stock.

```{r}
years <- c(2013:2016)
path_to_weightsdfm_list <- "weightsdfm_list_"
weightsdfm_list_type <- ".RData"
path_to_indices_list <- "indices_list_"
indices_list_type <- ".RData"

#TODO: unmap this cancer function

returns_quantiler <- function(sentiment_list, index_list, return_df, n_quantiles){
  sentiment_list %>%
  map(ntile, n = n_quantiles) %>%
  map(as.factor) %>%
  map(~ data_frame("quantile" = .)) %>%
  map2(index_list, cbind) %>%
  map(left_join, y = tickers, by = "filing") %>%
  map(left_join, y = return_df, by = "ticker") %>%
  map(group_by, quantile) %>%
  map(summarise, average_return = mean(return, na.rm = TRUE)) %>%
  map(transmute, relative_return = average_return / min(average_return) - 1) %>% # quantile 1 is row 1, 2 is row 2, etc. Higher quantile means higher sentiment value.
  do.call(what = cbind)
}

sentiment_returns_algo <- function(year, sentiment_dict){
load(paste(path_to_weightsdfm_list, year, weightsdfm_list_type, sep = "")) # loads into global
load(paste(path_to_indices_list, year, indices_list_type, sep = "")) # loads into global
sentiment_list <- weightsdfm_list %>%
  map(dfmstat_ratio, dict = sentiment_dict)
# names(returns_by_quantile) <- paste("section", sections, sep = "")
return(sentiment_list)
}

load("weightsdfm_list_2013.RData")

metadata <- map(paste0("metadata", years, ".csv"), read.csv) %>%
  reduce(rbind)

negative_sentiment <- map(years, sentiment_returns_algo, sentiment_dict = negative) %>%
  flatten %>%
  reduce(append)

#TODO: clean up this cancer code

names <- negative_sentiment %>% names %>%
  str_extract(pattern = ".*?(?=\\.)")

dummy <- data_frame(subsetted.names = names, negative_sentiment = negative_sentiment)
joined <- left_join(dummy, metadata, by = "subsetted.names") %>%
  group_by(filing, section) %>%
  filter(word.count == max(word.count)) %>%
  distinct(filing, section, .keep_all = TRUE) %>%
  select(filing, section, negative_sentiment) %>%
  spread(key = section, value = negative_sentiment) %>%
  rename(sec1sent = `1`, sec1Asent = `1A`, sec1Bsent = `1B`, sec2sent = `2`, sec3sent = `3`, sec4sent = `4`, sec5sent = `5`, sec6sent = `6`, sec7sent = `7`, sec7Asent = `7A`, sec8sent = `8`, sec9sent = `9`, sec9Asent = `9`, sec10sent = `10`, sec11sent = `11`, sec12sent = `12`, sec13sent = `13`, sec14sent = `14`, sec15sent = `15`)

masterIndex <- left_join(masterIndex, joined, by = "filing")
masterIndex <- read.csv("masterIndex.csv")

a <- spread(joined, key = section, value = negative_sentiment)

joined2 <- group_by(joined, filing, section) %>%
+ filter(word.count == max(word.count))

positive_returns_by_quantile <- map(years, sentiment_returns_algo, sentiment_dict = positive) %>%
  reduce(`+`) / length(years)
```
Load all the metadata together and rbind
Load all the names from the negative_sentiment and then perform a join operation.
Things in the metadata that are not in sentiment scores are not actual sections; e.g. they have sections tagged as section 16, 2014, 4A, etc.

Make the ggplot graph.

```{r}
quantile <- c(1:5) # little hacky but we need x to be the five quantiles; a proper melt would be the correct method, I think, but i cant get it to work. The problem is that this relies on the ggplot mapping 1 to quantile 1, 2 to quantile 2, etc. which works for c(1:5) but does not work for c("one", "two", "three", "four", "five"), for example, which will sort the character vector alphabetically

nm <- names(negative_returns_by_quantile)
negative_sentiment_quantile_returns <- map(nm, ~ ggplot(data = negative_returns_by_quantile, aes_string(x = quantile, y = .)) +
      geom_bar(stat = "identity") +
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())) %>%
  arrangeGrob(grobs = ., ncol = 5)
ggsave(negative_sentiment_quantile_returns, file = "negative_sentiment_quantile_returns.png")

nm <- names(positive_returns_by_quantile)
positive_sentiment_quantile_returns <- map(nm, ~ ggplot(data = positive_returns_by_quantile, aes_string(x = quantile, y = .)) +
      geom_bar(stat = "identity") +
      theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank())) %>%
  arrangeGrob(grobs = ., ncol = 5)
ggsave(positive_sentiment_quantile_returns, file = "positive_sentiment_quantile_returns.png")
```




```{r}
#this is test code and should be ignored

returns_by_quantile <- sentiment_list %>%
  map(ntile, n = 5) %>%
  map(as.factor) %>%
  map(~ data_frame("quantile" = .)) %>%
  map2(indices_list, cbind) %>%
  map(left_join, y = tickers, by = "filing") %>%
  map(left_join, y = return_df, by = "ticker") %>%
  map(group_by, quantile) %>%
  map(summarise, average_return = mean(return, na.rm = TRUE)) %>%
  map(transmute, relative_return = average_return / min(average_return) - 1) %>%
  do.call(what = cbind)
names(returns_by_quantile) <- rep(paste("section", sections))

quantiles <- map2(indices_list, quantile_negative, cbind)
returns_by_quantile <- map(quantiles, group_by, return) %>%
  map(summarise, average_return = mean(X2012, na.rm = TRUE))
```
